### 2012-10-13 Docs/Thesis/system_design_and_interface.tex
Thesis: evaluation interface description

\subsection{Efficient and Push-Based Evaluation}
\label{subsection:System_Design_and_Interface-Goals-Efficient_and_Push_based_Evaluation}

Efficient evaluation is the for push-based evaluation. Since FRP programs are
expected to  interact with an outside world in real time, efficiency cannot be
measured by runtime. Thus, when speaking of efficiency, we are expressing a
desire that the system utilize as few system resources as possible for the task
at hand, while responding as quickly as possible to external inputs and
producing output at a consistently high sample rate.

@@

We would like for the FRP system to be easily to integrate with such IO systems,
while being flexible enough to enable its use with other forms of IO systems,
such as simple imperative systems, threaded systems, or network servers.

@@

\subsection{Initialization and Running}

\subsection{Input and Output}

@@

>>>>

\subsection{Efficient Evaluation}
\label{subsection:System_Design_and_Interface-Goals-Efficient_and_Push_based_Evaluation}

Efficient evaluation is the motivation for push-based evaluation of events.
Since FRP programs are expected to  interact with an outside world in real time,
efficiency cannot be measured by the runtime of a program. Thus, when speaking of efficiency,
we are expressing a desire that the system utilize as few system resources as possible
for the task at hand, while responding as quickly as possible to external inputs and
producing output at a consistently high sample rate.

@@

We would like for the FRP system to be easy to integrate with such IO systems,
while being flexible enough to enable its use with other forms of IO systems,
such as simple imperative systems, threaded systems, or network servers.

@@

The evaluation interface provides a modified state monad which holds a signal
function, together with some additional information, as its state. Rather than
monadic instructions to put and get the state, the monad provides instructions
to trigger an input event, update an input signal, and trigger sampling of
signals in the signal function. Additional state includes the current set of
modifications to the input signals (since the last sample) and a set of
handlers which actuate effects based on output events or changes to the output
signal. The interface combinators are shown in Figure \ref{figure:evaluation_combinators}.

\begin{figure}
\begin{code}
data SVSample sv

data SVEventInput sv
instance SVRoutable SVEventInput sv

data SVSignalUpdate sv
instance SVRoutable SVSignalUpdate sv

data SVHandler out sv

svLeft          :: SVRoutable r =>
                   r svLeft -> r (svLeft :^: svRight)
svRight         :: SVRoutable r =>
                   r svRight -> r (svLeft :^: svRight)

svOcc           :: a -> SVEventInput (SVEvent a)
svSig           :: a -> SVSignalUpdate (SVSignal a)

emptyHandler    :: SVHandler out SVEmpty
signalHandler   :: (a -> out) -> SVHandler out (SVSignal a)
eventHandler    :: (a -> out) -> SVHandler out (SVEvent a)
combineHandlers ::    SVHandler out svLeft
                   -> SVHandler out svRight
                   -> SVHandler out (svLeft :^: svRight)

sample          :: a -> SVSample (SVSignal a)
sampleEvt       :: SVSample (SVEvent a)
sampleNothing   :: SVSample SVEmpty
combineSamples  ::    SVSample svLeft
                   -> SVSample svRight
                   -> SVSample (svLeft :^: svRight)

data SFEvalState m svIn svOut
initSFEval ::    SVHandler (m ()) svOut
              -> SVSample svIn
              -> Double
              -> (svIn :~> svOut)
              -> SFEvalState m svIn svOut

data SFEvalT svIn svOut m a
instance MonadTrans (SFEvalT svIn svOut)
instance (Monad m) => Monad (SFEvalT svIn svOut m)
instance (Functor m) => Functor (SFEvalT svIn svOut m)
instance (Monad m, Functor m) => Applicative (SFEvalT svIn svOut m)
instance (MonadIO m) => MonadIO (SVEvalT svIn svOut m)

runSFEvalT ::    SFEvalT svIn svOut m a
              -> SFEvalState m svIn svOut
              -> m (a, SFEvalState m svIn svOut)

push :: (Monad m) => SVEventInput svIn -> SFEvalT svIn svOut m ()
update :: (Monad m) => SVEventInput svIn -> SFEvalT svIn svOut m ()
step :: (Monad m) => Double -> SFEvalT svIn svOut m ()
\end{code}
\hrule
\caption{Evaluation combinators.}
\label{figure:evaluation_combinators}
\end{figure}

\subsection{Initialization and Running}
To evaluate a signal function, one first constructs an initial state
with the {\tt initSFEval} function. This function takes a set of
handlers, a set of initial samples for the signal inputs, the original
time (as a {\tt Double}) and the signal function to evaluate.

The signal function replaces itself after each event occurrence or time
step. The current signal function is maintained in the state. Calling
{\tt push} causes the signal function to react to the supplied event 
occurrence, and events which subsequently occur on the signal function's
output are passed to the supplied handlers. Calling {\tt update} updates
a set of new signal inputs. This set is nullified after it is passed in
along with a new time by calling {\tt step}. This allows various
components of a system to update individual signal inputs in a manner which is
decoupled from the sample timer.

When a monadic evaluation action is passed to the {\tt runSFEvalT} function,
along with a state, the result is an action in the underlying monad. This
action, when run in the underlying monad ({\tt IO}, for example) produces
a pair of an output value and a new state. This new state is what should be 
supplied to the next invocation of {\tt runSFEvalT}.

@@

{\tt SFEvalT} is not only a monad, but a monad transformer. This means that
it is parameterized by another monad, and that "running" an {\tt SFEvalT}
action produces an action in this underlying monad. It means also that actions
from this underlying monad may be ``lifted'' into actions of {\tt SFEvalT}.
So if the {\tt IO} monad is the underlying monad, {\tt IO} actions can be 
used to obtain input and arrange for timing. The underlying monad is also the
monad whose actions the output handlers produce, which provides the means of
actuating the output of the signal function.
